Formalizing Stack Safety as a Security Property

### 2021-07-08 submission3/main.tex
Fix Azevedo again; add testing to comparison with StkTokens

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
%\apt{maybe: ``Stack safety is a testable security property''?}\bcp{That
%  would sacrifice a lot of oomph for a little bit of precision, IMO.  Prefer
%  to leave it out of the title but hit it strongly in the abstract.}%
What does ``stack safety'' mean, exactly? The phrase is associated with a
variety of compiler,
run-time, and hardware mechanisms for protecting stack
memory, but these mechanisms typically lack precise specifications,
relying instead on informal descriptions and examples of the bad
behaviors that they prevent.

We propose a \apt{``generic''?} formal characterization
of stack safety based on concepts from language-based security: a
combination of an integrity property (``the private
state in each caller's stack frame is held invariant by the callee''),
a confidentiality property (``the callee's behavior is insensitive to the
caller's private state''), and a well-bracketedness property (``each
callee returns control to its immediate caller''). \apt{Do we really
  want to include WBCF in the abstract?} 
% These properties are sensible even in the presence of
% abnormal control flow.\apt{Is this important enough to mention here?}
% BCP: I think not in the abstract, but maybe in the intro.
% We further apply\apt{what does this mean? we didn't
%  invent it} a straightforward notion of
% ``well-bracketed control flow'' that restricts control flow to the
% typical pattern of matched calls and returns.

We use these properties to validate \apt{``evaluate'' ?} the stack-safety {\em micro-policies}
proposed by~\citet{DBLP:conf/sp/RoesslerD18}.  Specifically, we check (with
property-based random testing) that Roessler and Dehon's ``eager''
micro-policy, which catches violations as early as possible, enforces a
simple ``stepwise'' variant of our properties and correctly detects several
broken variants\apt{last phrase is grammatically disconnected}, and that (a repaired version of) their more performant
``lazy'' micro-policy corresponds to\apt{``enforces''?} a slightly weaker and more extensional
observational variant of our properties.
%
% \BCP{I still don't think the next sentence really nails it.  And (sigh) I am
%   still not really clear yet on whether our definition applies / might apply
%   to other enforcement mechanisms or not.  Saying that our attacker model is
%   stronger does not help---that should make it easier for us to describe
%   their protections, not harder.  So I don't understand what we're saying
%   here.}%
% %
% While we do not rule out other enforcement mechanisms, our
% definitions are strongly informed by micropolicy systems, especially
% our attacker model, which is stronger than other enforcement mechanisms
% can enable.

@@

%\bcp{Does use of unitialized variables lead to stack safety attacks?\apt{sure, via buffer overflow}}
%with the notion usually being
%defined negatively: through examples of stack-based exploits
%\rb{usually though not always; as noted below there are other
%  approaches}\bcp{Yes, we have to be very careful about this}.
%\leo{Examples at a high level}: such as---does not overwrite return addresses, or ...
\apt{This makes us seem very reactive to Skst.  Can we reorder things to give our story first, and
  compare to Skst afterwards?}
The most abstract characterization of stack safety that we are aware of is
recent work by \citet{SkorstengaardSTKJFP}, which defines stack safety as the conjunction of two
properties:
{\em local state encapsulation (LSE)} and {\em well-bracketed control flow (WBCF)}.
Informally, LSE says that the contents of
a caller's stack frame are not read or written while its callees (and their callees, etc.) are
executing, and WBCF says that callees always
return to the instruction following their call (if they return at all).
Skorstengaard et al.{} formalize these properties by defining an idealized
machine in which they hold {\em by construction};
they then show that an enforcement
mechanism (in their case, hardware capabilities) guarantees these properties, by
proving that an ordinary machine with enforcement simulates the idealized
machine.
%an enforcement mechanism (in their case, hardware
%capabilities)
%can then be shown to guarantee stack safety by exhibiting a fully abstract
%embedding function
%from the idealized machine to an ordinary machine equipped with this
%enforcement mechanism.  \bcp{Many OOPSLA readers/reviewers will not
%  understand that sentence.}
%%
%% While this work represents an important step in the right direction, we believe
%% there is still room to craft an operational characterization
%% of stack safety. \apt{Perhaps our rhetoric should be more micro-policy-specific here?}
%% In other words, we seek formal criteria that can be applied
%% to a mechanism for protecting stack-based low-level programs to judge
%% whether or not it actually ``guarantees stack safety.''
%% \rb{The remainder of this paragraph may be diving too quickly, would a reader be
%% confused by the references to these types of properties, yet to be introduced?}
%% We begin with a strong
%% local state encapsulation property that closely models the behavior of eager
%% micro-policies from \citet{DBLP:conf/sp/RoesslerD18}. This ``eager'' property
%% is too strong to characterize stack safety; reasonable enforcement mechanisms
%% exist that violate it. One such mechanism is a ``lazy'' tag policy, similar to
%% the one in \citet{DBLP:conf/sp/RoesslerD18}. We weaken local state encapsulation
%% to support this technique and argue that the new property, defined relative to an
%% arbitrary notion of observation. This property is still strong enough to prevent
%% the bad behaviors we associate with unsafe stack behavior, but weak enough that
%% other reasonable enforcement mechanisms can enforce it.

In the present work, we also describe the protection of both stack data (like LSE) and
control flow (like WBCF).  Our technical formulation of WBCF is similar to that of
Skorstengaard et al., but our treatment of stack data protection is rather different,
as is our approach to the definition itself. Instead of defining stack safety implicitly,
as ``what a certain idealized machine does,'' we give an
\emph{explicit} definition using the technical framework of language-based
security~\citep{sabelfeld2003language}, phrased as the conjunction of
\emph{stack integrity} and \emph{stack confidentiality}.
%
A key technical novelty in these definitions,
compared to standard formulations of confidentiality and integrity from the
security literature, is that they
are ``nested'': {\em each} caller is guaranteed protection from its
immediate callees (which, in turn, need protection from their immediate
callees, etc.). They are also stronger than LSE; even in the absence of the
control flow guarantees of WBCF, they guarantee callers'
data safety in the face of control-flow attacks (see \cref{sec:controlflow}).
%\rb{Reference to \cref{sec:running-example}, where this is discussed now?}
This reflects an arguably cleaner division between data protection and
control-flow protection than LSE / WBCF.

@@

To demonstrate the utility of our formal characterization, we use these
properties to validate and improve an existing enforcement mechanism, the
{\em stack-safety micro-policies} of~\citet{DBLP:conf/sp/RoesslerD18}, re-implemented
in the Coq proof assistant on top of a RISC-V specification.  We
use QuickChick~\citep{Denes:VSL2014,Pierce:SF4}, a property-based testing
tool for Coq, to generate random programs and check
that Roessler and Dehon's micro-policies correctly abort the ones that
attempt to violate one of our properties. Furthermore, we
%
check that the testing framework is able to generate counterexamples
that violate our properties but are \emph{not} halted by incorrect
enforcement variants---both variants that we accidentally created
during our re-implementation of the micro-policy and ones that we
intentionally crafted to be broken in order to increase our confidence
in testing and the enforcement mechanism itself.
%introduce several
%intentionally broken micro-policies and check that the testing framework is
%able to generate counterexamples that violate our properties but are not
%halted by the incorrect enforcement variants.  (Showing that the testing
%framework detects intentional bugs in enforcement mechanisms increases our
%confidence that it can also catch unknown ones, strengthening our confidence
%in the enforcement mechanism itself.\ifaftersubmission\bcp{Maybe that
%  sentence belongs in the testing section?}\fi)

We find that Roessler and Dehon's {\em Depth Isolation} micro-policy, in
which memory cells within each stack frame are tagged with the identity\apt{``depth''?} of
the function activation that owns the frame and access to those locations is
then permitted only when that activation\apt{``an actiation at that depth''?} is currently executing, validates our
stepwise properties. On the other hand, we reason\apt{but now we test?}  that \emph{Lazy Tagging and Clearing}
violates the temporal aspect of confidentiality in
corner cases where data can leak across repeated calls to the same callee,
and also violates integrity if the leak happens to use the caller's frame. We
propose a variant of {\em Lazy Tagging and Clearing} that \emph{should} enforce
confidentiality, albeit at some performance cost.\apt{Do we test?}
%
\ifaftersubmission
\bcp{Did Nick and Andre know about this deficiency?  Did they mention
  it in their paper?  (If neither, we should make a bigger deal out of the
  fact that we discovered it.)\apt{Their paper is a little vague on this subject.
    They refer to ``temporal security properties'' as something to do with dangling
    pointers, but they don't define it.
    Then they say that Depth Isolation has weaker temporal security
    guarantees than Static Authorities (tag=function name), but they don't explain why
    (and I don't see why). They don't say anything explicit about the temporal
    security properties of Lazy Tagging and Clearing, but they consider it to enforce
    a completely different (and generally weaker) policy anyway (``data flow
    integrity'' vs. ``memory safety'').}}\bcp{Sigh.  OK, I propose that we
  leave this issue to deal with later (in consultation with them, which we
  don't have time to do a good job of today).  At any rate, by my reading we
are not saying anything incorrect at the moment.}
\fi

@@

Lastly, we define integrity and confidentiality properties for a coroutine
system with a static layout, in which each coroutine's stack is constrained
to a fixed region of memory.  Though simplistic, this extension is a step
toward full-blown multithreading and demonstrates that our approach is flexible
enough to handle more sophisticated styles of control flow.

@@

%% %\bcp{I worry that we are being a little
%% %  dishonest here: the stated advantage of the lazy policy is that it is more
%% %efficient, but our variant is {\em not} efficient!}
In the rest of the paper, we begin with an informal overview of our key ideas via
examples (\cref{sec:running-example}), followed by a description of our key
assumptions about the machine environment and threat model (\cref{sec:setup}). 
We then offer the following contributions:

\begin{itemize}
\item We adapt concepts of integrity and confidentiality from language-based
security, combine these with an existing notion of well-bracketed
control flow to formalize a {\em stepwise stack-safety}
property, and argue that this property captures the intuitive concept of
strict stack
safety~(\cref{sec:prelim,sec:lse}).
\item We use property-based random testing to check that this stepwise
property is actually enforced by the {\em Depth Isolation} policy
of~\citet{DBLP:conf/sp/RoesslerD18}; in the process, we extend the generation by
execution techniques of \citet{TestingNI:ICFP} to handle nested
properties (\cref{sec:enforcement}).
%% \bcp{mention that we extend the ICFP 13 gen-by-exec idea with nested stuff...}
\item We formalize a
weaker but more extensional \emph{observational} variant of the stack safety
property that characterizes a refinement of Roessler and DeHon's more
performant {\em Lazy Tagging and  Clearing} policy (\cref{sec:lazy}).\apt{can this be more precise?}
\item We show that the definition can be
extended to argument passing on the stack, callee-saves registers, and a simple coroutine system (\cref{sec:ext}).
\end{itemize}

@@

Figure~\ref{fig:int1} shows a program with a simple integrity violation. It consists of a {\tt main}
function that performs a call to a function {\tt f}, which overwrites {\tt main}'s
variable {\tt x}. (Of course, this is illegal C, because the write in {\tt f} triggers
undefined behavior, but it is characteristic of what buggy or malicious code might do.) 
In the initial memory layout the entire stack is {\em unsealed},
meaning that it is eligible to be written to. At the call from {\tt main} to
{\tt f}, {\tt main}'s locals are {\em sealed} with {\tt main}'s depth,
in this case 0. Sealing is the fundamental integrity contract between the caller and callee:
{\tt main} expects that {\tt x} will be unchanged when it gets control back.
This gives us our simple, intuitive statement of {\em stepwise integrity}:
if a component is sealed before a step, it will be unchanged afterward. Stepwise
refers to the fact that the property must hold on every step. %Later we will
%loosen this condition in the interest of extensionality---the caller only
%cares about changes it can see after the return, not changes during execution.
In this case, after instruction 11, {\tt f} has violated integrity by writing to {\tt x},
so the final return value produced by {\tt main} is wrong.
%
(The black rectangle represents the return address, which is saved by the call
instruction itself, so it is not sealed.)

@@

Dually, confidentiality intuitively states that a callee is insensitive to
the context of its call (except for its arguments and any global
variables).
We state this as a kind of {\em noninterference} property~\citep{Goguen1982SecurityPA}.
Noninterference describes knowledge: it states that for system components
containing secret values, we can substitute any values whatsoever without changing the results
of our computation. If there exists some value for which we see different results, then
in fact we can learn something about the initial state by observing those results, violating
confidentiality.

In Figure~\ref{fig:conf1} we see that the callee, {\tt f}, violates confidentiality
by reading {\tt main}'s variable {\tt x}. The trace
shows how noninterference applies to this scenario. At instruction 10, we create a {\em variant}
state with arbitrary values\ifaftersubmission\bcp{IMO, it would be easier to
understand if we chose concrete values for the examples}\fi~\(v_0\), etc., in place of the values that should
be hidden from {\tt f}. (We show the evolving stack memory  of the variant in a second parallel column.) At instruction 12, we note that the callee's first variable, {\tt y}, has
changed to 5 in both the primary trace and the variant. The primary and variant agree on
the value in the changed location, so confidentiality was not violated. But at instruction 13,
the other variable, {\tt z}, has changed. In the primary trace it changed to 42, but
in the variant trace it might have any value \(v_0\). If \(v_0 \neq 42\) then the variant
has behaved differently than the primary, violating confidentiality.

@@

%From the classic stack smashing attack to return oriented programming,
One of the most pernicious ways to attack the stack is by hijacking control flow.
Our stack safety properties do not assume that control flow follows a typical structure
of matching sets of calls and returns nested within on another.
Rather, they aim to capture what it means to protect data even in the presence of control-flow attacks.
Consider Figure~\ref{fig:controlflow}.
(Here the diagram shows return addresses explicitly as line numbers, in the form $\#n$.)
Function {\tt f} is called twice, at instructions 2 and 4; it should therefore return to instructions
3 and 5, respectively. But the first time, before returning, it stashes its
return pointer away, and the second time it returns to instruction 3 instead of 5; then it ends
up being called a third time, at which point it finally returns to instruction 5.\apt{The line numbers are wrong.}

@@

In the example, on its first return, {\tt f} reaches its return target, because the
\(\PCname\) is 3 after its return. But {\tt f}'s second return is to the wrong instruction
(3 rather than 5) because the stack pointer has been overwritten (shown in light red).
Thus {\tt x} is still sealed and {\tt f}'s attempt to overwrite it at instruction 4
violates the integrity property.\apt{Ditto}

@@

Before introducing the model\apt{what does this really mean?}  itself, we set the context by stating some
assumptions we are making about machines, enforcement mechanisms, and
potential threats.

\paragraph*{Machine model}
Our properties are defined in terms of an abstract machine model that makes very few
assumptions about the concrete machine that instantiates it. Without changing the
properties, we can support\apt{what does this really mean?} any register
machine with a program counter (\(\PCname\)) and a stack pointer (\(\rsp\)); other
details of the machine's ISA are unimportant. We can support any static memory,
including a heap---our properties define stack protection regardless of how
non-stack memory is used. Likewise, we can support a range of calling convention
choices, like whether or not to have a frame pointer.

Some calling convention choices, such as callee-saves registers and argument
passing on the stack, require natural enhancements to our base properties.\apt{
  At this point it is unclear what these ``enhancements'' might be.}
We
present these enhancements, as well as a simple coroutine model,
in Section~\ref{sec:ext}, after we formalize the essence of stack safety.
Other plausible enhancements include:\apt{Better to move the ``what would
  be needed'' parts to the future work section -- they are mysterious here.}

@@

Runtime code generation would be much more challenging to support, as
it requires dynamic generation of code annotations. Similarly, true concurrency
would require more substantial changes to the machine
model. We consider these features out of scope for now.

\paragraph*{What is a ``call''?}  One significant challenge in formulating our
properties arises from the fact that we want to define stack safety at the
machine-code level for a variety of ISAs, including RISCy
\ifaftersubmission\sna{RISCy?}\leo{RISC-like maybe?}\sna{Apparently riscy is a thing.}\fi ones.  In
particular, there may be no single ``call'' instruction that does all the
required manipulation of the \(\PCname\) and stack pointer. Instead, there
might be just a simple branch instruction that stores the old \(\PCname\) in
a register, with the rest of the call sequence performed by other
instructions either before or after the jump.  Similarly, there may be no
``return'' instruction, but rather a ``return sequence'' involving an
indirect branch at some point.  Thus, the same instruction opcode may be
used in some places as part of a call or return sequence and in other places
for other purposes.

Since our stack-safety properties are all about protecting ``callers'' from
their ``callees,'' we need some clear-cut way to recognize when the program
makes or returns from a call.  Accordingly, we assume we are given a machine
language program {\em annotated} with the locations of instructions
representing calls.  We declare that a call has occurred when the machine
has just executed an instruction marked as a call.
%
These call annotations would typically come from a compiler together with
the machine code, but we do not assume anything about their provenance.
Annotations connect the enforcement mechanism to the property: we don't
protect unmarked calls, because a mechanism's correct functioning might
depend on knowing where calls are.

Returns, on the other hand, need not be annotated: we can tell when a callee
has returned control to the caller when...we see control return to the
caller! We don't require the callee's cooperation at all.
For example, under a standard calling convention, we can see that a
return has occurred when state of the machine reaches a state where
the stack pointer has the same value as it did at the point of some earlier
call and the $\PCname$ points to the instruction following that call.  We
abstract away from the details of specific calling conventions by assuming
that we are given some way of telling when a given machine state is a
{\em return target} with respect to some set of pending calls.

@@

\paragraph*{Enforcement Model}
\apt{What are our ambititions for genericity over enforcement?}

We model the base machine's behavior as a step function over machine states,
and a machine
enhanced with an enforcement mechanism, or {\em policy} for short, as a
partial step function over machine states extended with some
auxiliary state, refining the base machine. When the policy-enhanced machine
does not step from a given state, this is a {\em failstop}, halting the program before it can perform an
action that would violate the security property;
our properties are therefore naturally \emph{termination insensitive.}
This model will be made more precise in \cref{sec:prelim}.

Dividing policy state from machine state helps us reason about
enforcement mechanisms that keep a significant amount of data separate from the
primary execution data to inform security. Obviously tag-based systems do this\apt{At this
  point, we barely know what tags are},
with tags and extra policy state separated from, and untouchable by, application code.
Another significant example is the shadow stack\apt{citations? what enforcement systems do we actually have in mind?}, where a second stack
mirrors the real one, and a return may only occur if both stacks agree on the
target address. Typically the shadow stack is isolated using a separate protection
mechanism.

The separation is smaller in protection mechanisms that interact more directly
with application code.\apt{Vague. Mention capabilities?}  Some mechanisms might exhibit failstop behavior, but have
no protected policy state; others might never even failstop. Then the policy-enhanced
machine is simply identical to the original. This is the case for software-based
techniques.\apt{Even if they use a shadow stack?}

>>>>

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
%\apt{maybe: ``Stack safety is a testable security property''?}\bcp{That
%  would sacrifice a lot of oomph for a little bit of precision, IMO.  Prefer
%  to leave it out of the title but hit it strongly in the abstract.}%
What does ``stack safety'' mean, exactly? The phrase is associated with a
variety of compiler,
run-time, and hardware mechanisms for protecting stack
memory, but these mechanisms typically lack precise specifications,
relying instead on informal descriptions and examples of bad
behaviors that they prevent.

We propose a formal characterization
of stack safety, formulated with concepts from language-based security: a
combination of an integrity property (``the private
state in each caller's stack frame is held invariant by the callee''),
a confidentiality property (``the callee's behavior is insensitive to the
caller's private state''), and a well-bracketedness property (``each
callee returns control to its immediate caller'').
% These properties are sensible even in the presence of
% abnormal control flow.\apt{Is this important enough to mention here?}
% BCP: I think not in the abstract, but maybe in the intro.
% We further apply\apt{what does this mean? we didn't
%  invent it} a straightforward notion of
% ``well-bracketed control flow'' that restricts control flow to the
% typical pattern of matched calls and returns.

We use these properties to validate the stack-safety {\em micro-policies}
proposed by~\citet{DBLP:conf/sp/RoesslerD18}.  Specifically, we check (with
property-based random testing) that Roessler and Dehon's ``eager''
micro-policy, which catches violations as early as possible, enforces a
simple ``stepwise'' variant of our properties and correctly detects several
broken variants, and that (a repaired version of) their more performant
``lazy'' micro-policy corresponds to a slightly weaker and more extensional
observational variant of our properties.
%
% \BCP{I still don't think the next sentence really nails it.  And (sigh) I am
%   still not really clear yet on whether our definition applies / might apply
%   to other enforcement mechanisms or not.  Saying that our attacker model is
%   stronger does not help---that should make it easier for us to describe
%   their protections, not harder.  So I don't understand what we're saying
%   here.}%
% %
% While we do not rule out other enforcement mechanisms, our
% definitions are strongly informed by micropolicy systems, especially
% our attacker model, which is stronger than other enforcement mechanisms
% can enable.

@@

%\bcp{Does use of unitialized variables lead to stack safety attacks?\apt{sure, via buffer overflow}}
%with the notion usually being
%defined negatively: through examples of stack-based exploits
%\rb{usually though not always; as noted below there are other
%  approaches}\bcp{Yes, we have to be very careful about this}.
%\leo{Examples at a high level}: such as---does not overwrite return addresses, or ...
The most abstract characterization of stack safety that we are aware of is
recent work by \citet{SkorstengaardSTKJFP}, which defines stack safety as the conjunction of two
properties:
{\em local state encapsulation (LSE)} and {\em well-bracketed control flow (WBCF)}.
Informally, LSE says that the contents of
a caller's stack frame are not read or written while its callees (and their callees, etc.) are
executing, and WBCF says that callees always
return to the instruction following their call (if they return at all).
Skorstengaard et al.{} formalize these properties by defining an idealized
machine in which they hold {\em by construction};
they then show that an enforcement
mechanism (in their case, hardware capabilities) guarantees these properties, by
proving that an ordinary machine with enforcement simulates the idealized
machine.
%an enforcement mechanism (in their case, hardware
%capabilities)
%can then be shown to guarantee stack safety by exhibiting a fully abstract
%embedding function
%from the idealized machine to an ordinary machine equipped with this
%enforcement mechanism.  \bcp{Many OOPSLA readers/reviewers will not
%  understand that sentence.}
%%
%% While this work represents an important step in the right direction, we believe
%% there is still room to craft an operational characterization
%% of stack safety. \apt{Perhaps our rhetoric should be more micro-policy-specific here?}
%% In other words, we seek formal criteria that can be applied
%% to a mechanism for protecting stack-based low-level programs to judge
%% whether or not it actually ``guarantees stack safety.''
%% \rb{The remainder of this paragraph may be diving too quickly, would a reader be
%% confused by the references to these types of properties, yet to be introduced?}
%% We begin with a strong
%% local state encapsulation property that closely models the behavior of eager
%% micro-policies from \citet{DBLP:conf/sp/RoesslerD18}. This ``eager'' property
%% is too strong to characterize stack safety; reasonable enforcement mechanisms
%% exist that violate it. One such mechanism is a ``lazy'' tag policy, similar to
%% the one in \citet{DBLP:conf/sp/RoesslerD18}. We weaken local state encapsulation
%% to support this technique and argue that the new property, defined relative to an
%% arbitrary notion of observation. This property is still strong enough to prevent
%% the bad behaviors we associate with unsafe stack behavior, but weak enough that
%% other reasonable enforcement mechanisms can enforce it.

In the present work, we also describe the protection of both stack data (like LSE) and
control flow (like WBCF).  Our technical formulation of WBCF is similar to
theirs, but our treatment of stack data protection is rather different,
as is our approach to the definition itself. Instead of defining stack safety implicitly,
as ``what a certain idealized machine does,'' we give an
\emph{explicit} definition using the technical framework of language-based
security\ifaftersubmission~\citep{??}\bcp{Maybe the Sabelfeld and Myers
  survey?}\fi, phrased as the conjunction of
\emph{stack integrity} and \emph{stack confidentiality}.
%
A key technical novelty in these definitions,
compared to standard formulations of confidentiality and integrity from the
security literature, is that they
are ``nested'': {\em each} caller is guaranteed protection from its
immediate callees (which, in turn, need protection from their immediate
callees, etc.). They are also stronger than LSE; even in the absence of the
control flow guarantees of WBCF, they guarantee callers'
data safety in the face of control-flow attacks (see \cref{sec:controlflow}).
%\rb{Reference to \cref{sec:running-example}, where this is discussed now?}
This reflects an arguably cleaner division between data protection and
control-flow protection than LSE / WBCF.

@@

To demonstrate the utility of our formal characterization, we use these
properties to validate and improve an existing enforcement mechanism, the
{\em stack-safety micro-policies} of~\citet{DBLP:conf/sp/RoesslerD18}, re-implemented
in the Coq proof assistant on top of a RISC-V specification.  We
use QuickChick~\citep{Denes:VSL2014,Pierce:SF4}, a property-based testing
tool for Coq, to generate random programs and check
that Roessler and Dehon's micro-policies correctly abort the ones that
attempt to violate one of our properties; furthermore, we
%
check that the testing framework is able to generate counterexamples
that violate our properties but are not halted by incorrect
enforcement variants (both variants that we accidentally created
during our re-implementation of the micro-policy and ones that we
intentionally crafted to be broken in order to increase our confidence
in testing and the enforcement mechanism itself).
%introduce several
%intentionally broken micro-policies and check that the testing framework is
%able to generate counterexamples that violate our properties but are not
%halted by the incorrect enforcement variants.  (Showing that the testing
%framework detects intentional bugs in enforcement mechanisms increases our
%confidence that it can also catch unknown ones, strengthening our confidence
%in the enforcement mechanism itself.\ifaftersubmission\bcp{Maybe that
%  sentence belongs in the testing section?}\fi)

We find that Roessler and Dehon's {\em Depth Isolation} micro-policy, in
which memory cells within each stack frame are tagged with the identity of
the function activation that owns the frame and access to those locations is
then permitted only when that activation is currently executing, validates our
stepwise properties. On the other hand, we reason that \emph{Lazy Tagging and Clearing}
violates the temporal aspect of confidentiality in
corner cases where data can leak across repeated calls to the same callee,
and also violates integrity if the leak happens to use the caller's frame. We
propose a variant of {\em Lazy Tagging and Clearing} that \emph{should} enforce
confidentiality, albeit at some performance cost.
%
\ifaftersubmission
\bcp{Did Nick and Andre know about this deficiency?  Did they mention
  it in their paper?  (If neither, we should make a bigger deal out of the
  fact that we discovered it.)\apt{Their paper is a little vague on this subject.
    They refer to ``temporal security properties'' as something to do with dangling
    pointers, but they don't define it.
    Then they say that Depth Isolation has weaker temporal security
    guarantees than Static Authorities (tag=function name), but they don't explain why
    (and I don't see why). They don't say anything explicit about the temporal
    security properties of Lazy Tagging and Clearing, but they consider it to enforce
    a completely different (and generally weaker) policy anyway (``data flow
    integrity'' vs. ``memory safety'').}}\bcp{Sigh.  OK, I propose that we
  leave this issue to deal with later (in consultation with them, which we
  don't have time to do a good job of today).  At any rate, by my reading we
are not saying anything incorrect at the moment.}
\fi

@@

Lastly, we define integrity and confidentiality properties for a coroutine
system with a static layout, in which each coroutine's stack is constrained
to a fixed region of memory.  Though simplistic, this extension is a step
toward full-blown concurrency and demonstrates that our approach is flexible
enough to handle more sophisticated styles of control flow.

@@

%% %\bcp{I worry that we are being a little
%% %  dishonest here: the stated advantage of the lazy policy is that it is more
%% %efficient, but our variant is {\em not} efficient!}
In the rest of the paper, we begin with an informal overview of our key ideas via
examples (\cref{sec:running-example}), followed by definitions of some key
concepts (\cref{sec:setup}) and a discussion of the threat
model we consider (\cref{sec:threat}). We then offer the following contributions:

\begin{itemize}
\item We adapt concepts of integrity and confidentiality from language-based
security, combine these with an existing notion of well-bracketed
control flow to formalize a {\em stepwise stack-safety}
property, and argue that this property captures the intuitive concept of
strict stack
safety~(\cref{sec:prelim,sec:lse}).
\item We use property-based random testing to check that this stepwise
property is actually enforced by the {\em Depth Isolation} policy
of~\citet{DBLP:conf/sp/RoesslerD18}, and extend the generation by
execution techniques of \citet{TestingNI:ICFP} to handle nested
properties (\cref{sec:enforcement}).
%% \bcp{mention that we extend the ICFP 13 gen-by-exec idea with nested stuff...}
\item We formalize a
weaker but more extensional \emph{observational} variant of the stack safety
property that characterizes a refinement of Roessler and DeHon's more
performant {\em Lazy Tagging and  Clearing} policy (\cref{sec:lazy}).
\item We show that the definition can be
extended to argument passing on the stack and to a simple coroutine system (\cref{sec:ext}).
\end{itemize}

@@

Figure~\ref{fig:int1} shows a program with a simple integrity violation. It consists of a {\tt main}
function that performs a call to a function {\tt f}, which overwrites {\tt main}'s
variable {\tt x}. In the initial memory layout the entire stack is {\em unsealed},
meaning that it is eligible to be written to. At the call from {\tt main} to
{\tt f}, {\tt main}'s locals are {\em sealed} with {\tt main}'s depth,
in this case 0. Sealing is the fundamental integrity contract between the caller and callee:
{\tt main} expects that {\tt x} will be unchanged when it gets control back.
This gives us our simple, intuitive statement of {\em stepwise integrity}:
if a component is sealed before a step, it will be unchanged afterward. Stepwise
refers to the fact that the property must hold on every step. %Later we will
%loosen this condition in the interest of extensionality---the caller only
%cares about changes it can see after the return, not changes during execution.
In this case, after instruction 11, {\tt f} has violated integrity by writing to {\tt x},
so the final return value produced by {\tt main} is wrong.
%
(The black rectangle represents the return address, which is saved by the call
instruction itself, so it is not sealed.)

@@

Meanwhile, confidentiality intuitively states that a callee is insensitive to
the context of its call (except for its arguments and any global
variables).
We state this as a kind of {\em noninterference} property~\citep{Goguen1982SecurityPA}.
Noninterference describes knowledge: it states that for system components
containing secret values, we can substitute any values whatsoever without changing the results
of our computation. If there exists some value for which we see different results, then
in fact we can learn something about the initial state by observing those results, violating
confidentiality.

In Figure~\ref{fig:conf1} we see that the callee, {\tt f}, violates confidentiality
by reading {\tt main}'s variable {\tt x}. The trace
shows how noninterference applies to this scenario. At instruction 10, we create a {\em variant}
state with arbitrary values\ifaftersubmission\bcp{IMO, it would be easier to
understand if we chose concrete values for the examples}\fi~\(v_0\), etc., in place of the values that should
be hidden from {\tt f}. At instruction 12, we note that the callee's first variable, {\tt y}, has
changed to 5 in both the primary trace and the variant. The primary and variant agree on
the value in the changed location, so confidentiality was not violated. But at instruction 13,
the other variable, {\tt z}, has changed. In the primary trace it changed to 42, but
in the variant trace it might have any value \(v_0\). If \(v_0 \neq 42\) then the variant
has behaved differently than the primary, violating confidentiality.

@@

%From the classic stack smashing attack to return oriented programming,
One of the most pernicious ways to attack the stack is by hijacking control flow.
Our stack safety properties do not assume that control flow follows a typical structure
of matching sets of calls and returns nested within on another.
Rather, they aim to capture what it means to protect data even in the presence of control-flow attacks.
Consider Figure~\ref{fig:controlflow}.
(Here the diagram shows return addresses explicitly as line numbers, in the form $\#n$.)
Function {\tt f} is called twice, at instructions 2 and 4; it should therefore return to instructions
3 and 5, respectively. But the first time, before returning, it stashes its
return pointer away, and the second time it returns to instruction 3 instead of 5; then it ends
up being called a third time, at which point it finally returns to instruction 5.

@@

In the example, on its first return, {\tt f} reaches its return target, because the
\(\PCname\) is 3 after its return. But {\tt f}'s second return is to the wrong instruction
(3 rather than 5) because the stack pointer has been overwritten (shown in light red).
Thus {\tt x} is still sealed and {\tt f}'s attempt to overwrite it at instruction 4
violates the integrity property.

@@

Before introducing the model itself, let us set the context by stating some
assumptions we are making about machines, enforcement mechanisms, and
potential threats.

\paragraph*{Machine model}
Our properties are defined in terms of an abstract machine model that makes very few
assumptions about the concrete machine that instantiates it. Without changing the
properties, we can support any register
machine with a program counter (\(\PCname\)) and a stack pointer (\(\rsp\)); other
details of the machine's ISA are unimportant. We can support any static memory,
including a heap---our properties define stack protection regardless of how
non-stack memory is used. Likewise, we can support a range of calling convention
choices, like whether or not to have a frame pointer.

Some calling convention choices, such as callee-save registers and argument
passing on the stack, require natural enhancements to our base properties. We
present these enhancements, as well as a simple coroutine model,
in Section~\ref{sec:ext}, after we formalize the essence of stack safety.
Other plausible enhancements include:

@@

Runtime code generation would be much more challenging to support, as
it requires dynamic code annotations. Similarly, parallelism beyond the current
coroutine model would require more substantial changes to the machine
model. We consider these features out of scope for now.

\paragraph*{What is a ``call''?}  One significant challenge in formulating our
properties arises from the fact that we want to define stack safety at the
machine-code level for a variety of ISAs, including RISCy
\ifaftersubmission\sna{RISCy?}\leo{RISC-like maybe?}\sna{Apparently riscy is a thing.}\fi ones.  In
particular, there may be no single ``call'' instruction that does all the
required manipulation of the \(\PCname\) and stack pointer. Instead, there
might be just a simple branch instruction that stores the old \(\PCname\) in
a register, with the rest of the call sequence performed by other
instructions either before or after the jump.  Similarly, there may be no
``return'' macro-instruction, but rather a ``return sequence'' involving an
indirect branch at some point.  Thus, the same instruction opcode may be
used in some places as part of a call or return sequence and in other places
for other purposes.

Since our stack-safety properties are all about protecting ``callers'' from
their ``callees,'' we need some clear-cut way to recognize when the program
makes or returns from a call.  Accordingly, we assume we are given a machine
language program {\em annotated} with the locations of instructions
representing calls.  We declare that a call has occurred when the machine
has just executed an instruction marked as a call.
%
These call annotations would typically come from a compiler together with
the machine code, but we do not assume anything about their provenance.
Annotations connect the enforcement mechanism to the property: we don't
protect unmarked calls, because a mechanism's starting conditions might
depend on knowing where calls are.

Returns, on the other hand, need not be annotated: we can tell when a callee
has returned control to the caller when... we see control return to the
caller! We don't require the callee's cooperation at all.
For example, under a standard calling convention, we can see that a
return has occurred when state of the machine reaches a state where
the stack pointer has the same value as it did at the point of some earlier
call and the $\PCname$ points to the instruction following that call.  We
abstract away from the details of specific calling conventions by assuming
that we are given some way of telling when a given machine state is a
{\em return target} with respect to some set of pending calls.

@@

\paragraph*{Enforcement Model}
We model the base machine's behavior as a step function over machine states,
and a machine
enhanced with an enforcement mechanism, or {\em policy} for short, as a
partial step function over machine states extended with some
auxiliary state, refining the base machine. When the policy-enhanced machine
does not step from a given state, this is a {\em failstop}, halting the program before it can perform an
action that would violate the security property;
our properties are therefore naturally \emph{termination insensitive.}
This model will be made more precise in \cref{sec:prelim}.

Dividing policy state from machine state helps us reason about
enforcement mechanisms that keep a significant amount of data separate from the
primary execution data to inform security. Obviously tag-based systems do this,
with tags and extra policy state separated from and untouchable by application code.
Another significant example is the shadow stack, where a second stack
mirrors the real one, and a return may only occur if both stacks agree on the
target address. Typically the shadow stack is isolated using a separate protection
mechanism.

The separation is smaller in protection mechanisms that interact more directly
with application code. Some mechanisms might exhibit failstop behavior, but have
no protected policy state; others might never even failstop. Then the policy-enhanced
machine is simply identical to the original. This is the case for software-based
techniques.
